Parity games in pushdown systems with parameters

### 2023-07-04 reachability/article.tex
Satisfy the 12 pages limit now

Once we have provided players with an arena, it remains to define  properly what the winning condition is in order to define a game. Given an arena $A = (S_0, S_1, \rightarrow)$, a {\em winning condition}
$\win$
is a subset  
of the set of maximal plays for $A$. A {\em game} is a pair $\mathcal{G}= (A, \win)$ which is composed of an arena $A$ and a winning condition $\win$.
Although a winning condition is in general an infinite object we are interested in winning conditions 
that can finitely be described.
A strategy $\sigma_0$ for player $0$ from position $s \in S$ is a {\em winning strategy
 from $s$} for player $0$ if 
every maximal play starting from $s$ and consistent with $\sigma_0$ is in $\win$.
On the other hand, a strategy $\sigma_1$ for player $1$ from position $s \in S$ is a {\em winning strategy
 from $s$} for player $1$ if 
every maximal play starting from $s$ and consistent with $\sigma_1$ is not in $\win$.

We consider {\em parity games}: 
the winning condition $\win_\Omega(S)$ is given by a {\em priority function} 
 $\Omega : S \to  [0, m]$.
An infinite path $ \pi = s_0 s_1 s_2 \ldots$ is in $\win_\Omega(S)$ if the smallest priority appearing infinitely often in
the sequence 
 is even, i.e.
if $\min( \{ \Omega(s) | ~ s \in \text{\textit{Inf}}( \pi) \})$ is even.  
A finite play is in $\win_\Omega(S)$ if its target configuration is in $S_1$.
We simply write
$\win_\Omega$ in case the set $S = S_0 \biguplus S_1$ is obvious from context.

@@

\par\noindent\ignorespacesafterend
A {\em parametric pushdown system} ({\em PPDS} for short) 
is a tuple $\mathcal{Z} = (Q, \Gamma, P, R, q_{init},\gamma_{init}, F),$ where
 $Q$ is a non-empty finite {\em set of  states};
 $\Gamma$ is a non-empty finite {\em  stack alphabet};
 $P$ is a finite {\em   set of parameters}, \textcolor{black}{with $\Gamma \cap P = \emptyset$};
  $R   \subseteq  Q  \times (\Gamma \biguplus P)  \times Q  \times \Op(\Gamma)$ is finite {\em  set of rules};
 $q_{init}\in Q$ is an {\em initial  state};
 $ \gamma_{init} \in \Gamma$ is an {\em initial stack symbol}; and
 $F\subseteq Q$ is a {\em set of final  states}.

@@

\begin{samepage}
A parametric pushdown automaton  $\mathcal{Z}= (Q, \Gamma, P,u R, q_{init}, \gamma_{init},F)$ 
and a parameter valuation $\mu: P\rightarrow \Gamma^*$ induce the 
transition system  
$T_{\mathcal{Z}}^{\mu} = (\Conf(\mathcal{Z}), 
\rightarrow_{\mathcal{Z},\mu})$ where
for all $ q, q'  \in  Q$, 
for all $ w,w' \in \Gamma^* $, and 
for all $ a  \in  \Gamma$, 
$q(wa) \rightarrow_{\mathcal{Z},\mu} q'(w')$
if there exists a rule in $R$
 of the form $(q,a,q', op)$ or $(q,p,q', op)$ with $\mu(p)=wa$,
such that
either of the following holds
\begin{itemize}
\item $ op = \text{\textit{push}}^\gamma$ and $w' = wa\gamma$,

@@

We introduce here arenas over the set of partial parameter valuations for a set of parameters
$P = P_0 \biguplus P_1$. The values are assigned to the parameters one by one, alternating between values assigned by player $0$ (corresponding to parameters in $P_0$) and by player $1$ (corresponding to parameters in $P_1$). Other options could have been considered: for instance, we could consider a fully existential approach 
(i.e. all parameters have their values assigned by player $0$), 
or a fully universal approach
(i.e. all parameters have their values assigned by player $1$). 
Note however that the approach considered here encompasses both of the latter approaches:
the fully existential approach consists in the case where $P_1 = \emptyset$ and
the fully universal one, the case $P_0 = \emptyset$.

Assume an ordering of the parameters $P_0 \biguplus P_1 = \{p_0, p_1, \ldots, p_k \}$ for some 
 $k \in \N$.
%
Then given a alphabet $\Gamma$, a {\em parameter valuation arena} is a tuple $A_{P_0,P_1,\Gamma} = (S_0, S_1, \rightarrow_P)$ where
\begin{itemize}

@@

Note that the arena has dead ends, and that these are the configurations that correspond to parameter valuations.
Given a strategy $\sigma_0$ for player $0$ and a strategy $\sigma_1 $ for player $1$,
the last configuration of the resulting play $\pi(\mu_\bot,\sigma_0, \sigma_1)$
is called {\em the resulting parameter valuation}
 and is denoted by $\mu_{\sigma_0, \sigma_1}$, where 
 $\mu_\bot$ is the partial parameter valuation with domain $\emptyset$. 
We then use parametric valuation arenas as prefixes in parametric games: once the resulting parameter valuation is produced by the two players, it serves to instantiate the transition system for continuing the game.
%
%
Indeed, where a concrete automaton induces a transition system, a parametric automaton induces one transition system for every possible parameter valuation.
We use $T^\mu_\A$ to denote the transition system induced by the concrete automaton corresponding to
a parametric automaton $\A$ with parameters $P$ taking values in $\Gamma^*$ and a parameter valuation $\mu: P \to \Gamma^*$.

More formally, for an  $n$-parametric pushdown automaton 
$\mathcal{Z}= (Q, \Gamma, P, R, q_{init}, \gamma_{init},F)$,
given a partition of
$Q$ into $Q_0$ and $Q_1$, 
we naturally partition the configurations of
 $\mathcal{Z}$
into
$\Conf_{\mathcal{Z},0}=Q_0\times \Gamma^*$
and
$\Conf_{\mathcal{Z},1}=Q_1\times \Gamma^*$. 
With these notations in mind we define the arena
$
A_{(\mathcal{Z}, Q_0, Q_1, \mu)}
=
(\Conf_{\mathcal{Z},0}, \Conf_{\mathcal{Z},1}, \rightarrow_{\mathcal{Z},\mu})$
induced by $\mathcal{Z}$, the partition $Q = Q_0 \biguplus Q_1$, and a
parameter valuation $\mu: P\rightarrow \Gamma^*$.
From a partition of
$P$ into $P_0$ and $P_1$, 
we thus define 
$A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}$
as
the arena that contains both the configurations and transitions of
the parameter valuation arena
$A_{P_0,P_1,\Gamma}$,
and that moreover contains the configurations and transitions of
$A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1, \mu)}$ 
\---- albeit with configurations additionally including $\mu$ in the tuple to avoid confusion \----
for
all parameter valuations $\mu: P\rightarrow \Gamma^*$.
Moreover, for every configuration $\mu$ in $A_{P_0,P_1,\Gamma^*}$,
we add a transition
from $\mu$ towards 
$q_{init}(\gamma_{init}, \mu)$ in $A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1, \mu)}$. \label{PPDA reachability game}

We extend any 
priority function $\Omega: Q \to [0, m]$
into
$\widehat{\Omega}: \Conf(\mathcal{Z}) \cup (\Gamma^* \cup \bot)^{[0,n]} \to [0, m]$,
where we set
$\widehat{\Omega}(q, w, \mu) = \Omega(q)$
for all
$w \in \Gamma^*$
and
for all $\mu \in (\Gamma^*)^{P}$,
and
$\widehat{\Omega}(\mu) = 0$
for all $\mu \in (\Gamma^* \cup \bot)^{P}$.

\problemx{Parametric pushdown parity game}
{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init}, \gamma_{init} , F)$, where $Q = Q_0  \biguplus Q_1 $, and $ P = P_0 \biguplus P_1$, 
and a priority function $\Omega: Q \to [0, m]$.}
{Does player $0$ have a winning strategy from $\mu_\bot$ in the parity game \\
$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{\widehat{\Omega}})$ ?\newline}

@@

A pebble automaton is a two-way finite state automaton that uses a fixed, finite number of
pebbles that it can drop on, and lift from 
words, using them as markers.
Pebble automata recognize regular languages only, provided the life times of the pebbles, i.e. the times between dropping a pebble and lifting it again, are properly nested
\cite{globerman1996complexity, engelfriet1999tree}.
Automata with nested pebbles were also introduced
for tree-walking automata. 
It is known that tree-walking automata do not recognize all regular tree languages \cite{bojanczyk2008tree}.
Using pebbles is a remedy against getting lost along a tree, but 
the unrestricted use of pebbles leads to a class of tree languages much larger than the
regular tree languages, in fact to all tree languages in $\NSPACE(log ~ n)$.
Thus, in  both pebble word automata and pebble tree-walking automata, the placement of the pebble follows a strict stack discipline. It is traditional hence to represent syntactically the dropping and lifting of pebbles by operations \textit{lift} and \textit{drop}; a \textit{drop} simply records the current position with a fresh pebble (such a pebble should be available)  
and a \textit{lift} pops the last dropped pebble if the current position corresponds to the one recorded by it.

@@

We first define more formally our pebble pushdown automaton framework, and then provide a reduction from the problem of solving parametric pushdown parity games to the problem of solving pebble pushdown parity games.

\subsection{Pebble pushdown systems}

@@

An {\em $n$-pebble pushdown automaton} 
is a tuple 
$\ppda = (Q, \Gamma,  R, q_{init}, \gamma_{init}, F)$
where $Q$ is a {\em finite set of  states}; $\Gamma$ is a {\em finite stack alphabet};
  $R  \subseteq  Q  \times \Gamma \times
		 \{ 0, 1, \ldots n\} \times \mathcal{P}(\{1,\ldots,n\})
		 \times Q  \times (\Op(\Gamma) \cup \{\text{\textit{drop}}, \text{\textit{lift}}\})$ is a {\em finite set of rules},
		 where the fourth element $S$ of a rule $r$ is a subset of 
		$\mathcal{P}(\{1,\ldots,i\})$ where $i$ is the third element of $r$,
		and if the last element is
		$\text{\textit{lift}}$ then we additionally require $i \in S$;
 $q_{init} \in Q$ is an {\em initial  state};
 $\gamma_{init} \in \Gamma$ is an {\em initial stack symbol}; and
 $F \subseteq Q$ is a {\em set of final  states}.

@@

The idea of a transition $(q, a, i, S, q', m) \in R$
is that, if the automaton $\ppda$ is in state $q$ with pebbles $1,\ldots, i$ dropped \---- or without pebble dropped if $i = 0$ \---- with top stack symbol $a$, and stack content which corresponds to the stack contents of the pebbles from $S$ and only these pebbles, then
$\ppda$ goes to state 
$q'$ and makes modifications to the stack or the pebbles according to
$m$. 
Note a pebble 
can be lifted only if the stack content which corresponds to the pebble
is the same as the current stack content. 
This is enforced by syntactically requiring
the last pebble dropped $i$ is in the set $S$ used for testing the presence of certain pebbles.

@@

\begin{samepage}
An $n$-pebble pushdown automaton $\ppda =  (Q, \Gamma,  R, q_{init}, {\gamma_{init}} , F)$ induces 
a transition system  $T_{\ppda} = (\Conf(\ppda), \rightarrow_{\ppda})$ 
 such that
for all $(q, a, i, S, q', m) \in R$, with $a \in \Gamma$,
 for all words $w \in  \Gamma^*$, and for all $\{ 1, \ldots , i \}$-pebble assignments $\mu$ such that $\mu(j) = wa$ 
 for each
$j \in S$ 	and
 $\mu(j) \neq wa$ 
 for each $j \in \{ 1, \ldots , i \} \setminus S$,
the following holds
\begin{itemize}
\item if $m \in \Op(\Gamma)$, 
either
\begin{itemize}
\item $ m = \text{\textit{push}}^\gamma$ and $(q,wa,\mu) \rightarrow_{\ppda} (q',wa\gamma, \mu)$,

@@

\item $ m = \text{\textit{skip}}$ and $(q,wa,\mu) \rightarrow_{\ppda} (q',wa, \mu)$,
\end{itemize}
\item if $ m = \text{\textit{drop}}$, 
$(q,wa,\mu) \rightarrow_{\ppda} (q',wa, \mu')$,  
where $\mu'$ is the $\{ 1, \ldots , i, i+1 \}$-pebble assignment such that
$\mu'(j) = \mu(j)$, for each
$j \leq i$, and $\mu'(i+1) = wa$, and
\item if $m = \text{\textit{lift}}$, and $i \in S$, i.e. the last pebble dropped belong of the set of pebble we test the presence of,
$(q,wa,\mu) \rightarrow_{\ppda} (q',wa, \mu')$, 
where $\mu'$ is the $\{ 1, \ldots , i - 1 \}$-pebble assignment such that
$\mu'(j) = \mu(j)$, for each
$j < i$.
\end{itemize}
\end{samepage}

\par\noindent\ignorespacesafterend
We are interested in games over pebble pushdown systems, 
mainly, parity games.

@@

\problemx{$n$-pebble pushdown parity game}
{An $n$-pebble pushdown automaton $\ppda= (Q, \Gamma,  R, q_{init}, {\gamma_{init}} , F)$, 
where $Q = Q_0  \biguplus Q_1 $, 
and
a priority mapping $\Omega: Q \to \{0, \ldots m\}$.}
{Does player $0$ have a winning strategy from $q_{init}(\gamma_{init}, \mu_{init})$ for the parity game 
$\mathcal{G}= (A_{(\ppda, Q_0 , Q_1)}, \win_{\overline{\Omega}})$ ?}

@@

\subsection{Higher-order pushdown systems}

@@

The following operations can be performed on a $1$-stacks of length $m \in \N$.
\begin{eqnarray*}
&\text{\textit{push}}_1^\gamma( w_0 w_1 \ldots w_m) &=\quad w_0 w_1 \ldots w_m \gamma   \text{ for all } \gamma \in  \Gamma, \\
&\text{\textit{pop}}_1(w_0 w_1 \ldots w_{m-1} w_m ) &=\quad w_0 w_1 \ldots w_{m-1}, \text{ if } m \geq 1\\
&\text{\textit{top}}( w_0 w_1 \ldots w_m ) &=\quad w_m.
\end{eqnarray*}
The operations added at level $n+1$ are the copy of the topmost $n$-stack
and the removal of the topmost $n$-stack.
More formally, if $ \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle$ is a stack of 
level $n > 1$, 
the following operations are possible.
%
\begin{eqnarray*}
&\text{\textit{push}}_n( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &=\quad  \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle \langle s_m \rangle , \\
&\text{\textit{push}}_j( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &=\quad \langle s_0 \rangle, \langle s_1 \rangle \ldots 
\langle \text{\textit{push}}_j(s_m) \rangle,
\text{ if } 2 \leq j < n\\
&\text{\textit{push}}_1^\gamma( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &= \quad
\langle s_0 \rangle \langle s_1 \rangle \ldots 
\langle \text{\textit{push}}_1^\gamma(s_m) \rangle,
\text{ for all } \gamma \in \Gamma,\\
&\text{\textit{pop}}_n( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle \langle s_m \rangle) &= \quad
\langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle
,\\
&\text{\textit{pop}}_j( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle \langle s_m \rangle) &= \quad
\langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle
\langle \text{\textit{pop}}_j(s_m) \rangle,
\text{ if } 1 \leq j < n\\
&\text{\textit{top}}( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &=\quad
\text{\textit{top}}(s_m).
\end{eqnarray*}
The operations $pop_1$ and $top$ are undefined on a stack whose top $1$-stack is empty.  \\

@@

\begin{samepage}
\par\noindent\ignorespacesafterend
A {\em higher-order pushdown systems of level $n$} (or $n$-HPDS for short) 
is a tuple $\H=(Q, \Gamma, R, q_{init}, \gamma_{init}, F)$,
where 
\begin{itemize}
        \item $Q$ is a non-empty finite {\em set of  states},
        \item $\Gamma$ is a non-empty finite {\em  stack alphabet},
	\item $R\subseteq Q\times \Gamma \times Q \times \Op_n(\Gamma)$ is a finite {\em set of  rules},
        \item $q_{init}$ is the {\em initial  state},
        \item $\gamma_{init} \in \Gamma$ is the {\em %bottom-of-stack
initial stack symbol}, and
	\item $F\subseteq Q$ is a {\em set of final  states}.
		\end{itemize}
\end{samepage}

\par\noindent\ignorespacesafterend
By $\Conf(\mathcal{H}) = Q \times \mathscr{S}_n(\Gamma)$
we denote the set of {\em configurations} of an $n$-HPDS $\H$. 
As usual we abbreviate $(q,s) \in \Conf(\mathcal{H})$ as $q(s)$.

An $n$-HPDS $\H=(Q, \Gamma, R, q_{init}, F)$
induces the transition system $\T_\H = (Conf(\mathcal{H}), \rightarrow_{\mathcal{H}})$
where
for all $q,q'$ in $Q$,
for all $s,s'$ in $\mathscr{S}_n(\Gamma)$,
$q(s) \rightarrow_{\mathcal{H}} q'(s')$ if
there exists some rule
$(q,\gamma,q',op) \in R$
such that
$\text{\textit{top}}(s) = \gamma$
and
$s' = op(s)$.

Again we are interested in parity games.
As expected, given an 
 $n$-HPDS $\mathcal{H}$
 and a partition of
$Q$ into $Q_0$ and $Q_1$,
we partition the configurations of
$T_{\mathcal{H}}$
into
$\Conf_{\mathcal{H},0}=Q_0\times \mathscr{S}_n(\Gamma)$
and
$\Conf_{\mathcal{H},1}=Q_1\times\mathscr{S}_n(\Gamma)$.
%
%
With these notations in mind one can define the arena
$$
A_{(\mathcal{H}, Q_0, Q_1)}
=
(\Conf_{\mathcal{H},0}, \Conf_{\mathcal{H},1}, \rightarrow_{\mathcal{H}})$$
\par\noindent\ignorespacesafterend
induced by an $n$-HPDS $\mathcal{H}$ and a partition of its set of states.

>>>>

Once we have provided players with an arena, it remains to define  properly what the winning condition is in order to define a game. Given an arena $A = (S_0, S_1, \rightarrow)$, a {\em winning condition}
$\win$
is a subset  
of the set of maximal plays for $A$. A {\em game} is a pair $\mathcal{G}= (A, \win)$ which is composed of an arena $A$ and a winning condition $\win$.
Although a winning condition is in general an infinite object we are interested in parity games winning conditions, 
which can finitely be described. In 
{\em parity games},
the winning condition $\win_\Omega(S)$ is given by a {\em priority function} 
 $\Omega : S \to  [0, m]$.
An infinite path $ \pi = s_0 s_1 s_2 \ldots$ is in $\win_\Omega(S)$ if the smallest priority appearing infinitely often in
the sequence 
 is even, i.e.
if $\min( \{ \Omega(s) | ~ s \in \text{\textit{Inf}}( \pi) \})$ is even.  
A finite play is in $\win_\Omega(S)$ if its target configuration is in $S_1$.
We simply write
$\win_\Omega$ in case the set $S = S_0 \biguplus S_1$ is obvious from context.

A strategy $\sigma_0$ for player $0$ from position $s \in S$ is a {\em winning strategy
 from $s$} for player $0$ if 
every maximal play starting from $s$ and consistent with $\sigma_0$ is in $\win$.
On the other hand, a strategy $\sigma_1$ for player $1$ from position $s \in S$ is a {\em winning strategy
 from $s$} for player $1$ if 
every maximal play starting from $s$ and consistent with $\sigma_1$ is not in $\win$.

@@

\par\noindent\ignorespacesafterend
A {\em parametric pushdown system} ({\em PPDS} for short) 
is a tuple $\mathcal{Z} = (Q, \Gamma, P, R, q_{init},\gamma_{init}),$ where
 $Q$ is a non-empty finite {\em set of  states};
 $\Gamma$ is a non-empty finite {\em  stack alphabet};
 $P$ is a finite {\em   set of parameters}, \textcolor{black}{with $\Gamma \cap P = \emptyset$};
  $R   \subseteq  Q  \times (\Gamma \biguplus P)  \times Q  \times \Op(\Gamma)$ is finite {\em  set of rules};
 $q_{init}\in Q$ is an {\em initial  state};
 $ \gamma_{init} \in \Gamma$ is an {\em initial stack symbol}.

@@

\begin{samepage}
A parametric pushdown automaton  $\mathcal{Z}= (Q, \Gamma, P,u R, q_{init}, \gamma_{init})$ 
and a parameter valuation $\mu: P\rightarrow \Gamma^*$ induce the 
transition system  
$T_{\mathcal{Z}}^{\mu} = (\Conf(\mathcal{Z}), 
\rightarrow_{\mathcal{Z},\mu})$ where
for all $ q, q'  \in  Q$, 
for all $ w,w' \in \Gamma^* $, and 
for all $ a  \in  \Gamma$, 
$q(wa) \rightarrow_{\mathcal{Z},\mu} q'(w')$
if there exists a rule in $R$
 of the form $(q,a,q', op)$ or $(q,p,q', op)$ with $\mu(p)=wa$,
such that
either of the following holds
\begin{itemize}
\item $ op = \text{\textit{push}}^\gamma$ and $w' = wa\gamma$,

@@

For an  $n$-parametric pushdown automaton 
$\mathcal{Z}= (Q, \Gamma, P, R, q_{init}, \gamma_{init})$,
and a partition
$P = P_0 \biguplus P_1$,
we introduce first arenas over the set of partial parameter valuations. The values are assigned to the parameters one by one, alternating between values assigned by player $0$ (corresponding to parameters in $P_0$) and by player $1$ (corresponding to parameters in $P_1$). Other options could have been considered: for instance, we could consider a fully existential approach 
(i.e. all parameters have their values assigned by player $0$), 
or a fully universal approach
(i.e. all parameters have their values assigned by player $1$). 
Note however that the approach considered here encompasses both of the latter approaches:
the fully existential approach consists in the case where $P_1 = \emptyset$ and
the fully universal one, the case $P_0 = \emptyset$.

Assume an ordering of the parameters $P_0 \biguplus P_1 = \{p_0, p_1, \ldots, p_k \}$ for some 
 $k \in \N$, the {\em parameter valuation arena} is a tuple $A_{P_0,P_1,\Gamma} = (S_0, S_1, \rightarrow_P)$ where
\begin{itemize}

@@

The arena has dead ends which correspond to parameter valuations.
Given a strategy $\sigma_0$ for player $0$ and a strategy $\sigma_1 $ for player $1$,
the last configuration of the resulting play $\pi(\mu_\bot,\sigma_0, \sigma_1)$
is called {\em the resulting parameter valuation}
 and is denoted by $\mu_{\sigma_0, \sigma_1}$, where 
 $\mu_\bot$ is the partial parameter valuation with domain $\emptyset$. 
The parametric valuation arena is a prefix of the full parametric games: once the resulting parameter valuation is produced by the two players, it serves to instantiate the transition system for continuing the game.
%
%

We then obtain the full parametric game's arena $A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}$ by, for every configuration $\mu$ in $A_{P_0,P_1,\Gamma^*}$,
 adding a transition
from $\mu$ towards the initial configuration
$(q_{init}(\gamma_{init}), \mu)$ 
of
the arena
$
A_{(\mathcal{Z}, Q_0, Q_1, \mu)}
=
(Q_0\times \Gamma^*, Q_1\times \Gamma^*, \rightarrow_{\mathcal{Z},\mu})$
with configurations additionally including $\mu$ in the tuple.

 \label{PPDA reachability game}

We extend any 
priority function $\Omega: Q \to [0, m]$
into
$\widehat{\Omega}: \Conf(\mathcal{Z}) \cup (\Gamma^* \cup \bot)^{[0,n]} \to [0, m]$,
where we set
$\widehat{\Omega}(q, w, \mu) = \Omega(q)$
for all
$w \in \Gamma^*$
and
for all $\mu \in (\Gamma^*)^{P}$,
and
$\widehat{\Omega}(\mu) = 0$
for all $\mu \in (\Gamma^* \cup \bot)^{P}$. We consider the following decision problem.

\problemx{Parametric pushdown parity game}
{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init}, \gamma_{init} )$, where $Q = Q_0  \biguplus Q_1 $, and $ P = P_0 \biguplus P_1$, 
and a priority function $\Omega: Q \to [0, m]$.}
{Does player $0$ have a winning strategy from $\mu_\bot$ in the parity game \\
$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{\widehat{\Omega}})$ ?\newline}

@@

A pebble automaton is a two-way finite state automaton that uses a fixed, finite number of
pebbles that it can drop on, and lift from 
words, using them as markers.
Pebble automata recognize regular languages only, provided the life times of the pebbles, i.e. the times between dropping a pebble and lifting it again, are properly nested
\cite{globerman1996complexity, engelfriet1999tree}.
Automata with nested pebbles were also introduced
for tree-walking automata. 
It is known that tree-walking automata do not recognize all regular tree languages \cite{bojanczyk2008tree}.
Using pebbles is a remedy against getting lost along a tree, but 
the unrestricted use of pebbles leads to a class of tree languages much larger than the
regular tree languages, in fact to all tree languages in $\NSPACE(log ~ n)$.
It is traditional since pebbles follows a strict stack discipline to represent syntactically the dropping and lifting of pebbles by operations \textit{lift} and \textit{drop}; a \textit{drop} simply records the current position with a fresh pebble (such a pebble should be available)  and a \textit{lift} pops the last dropped pebble if the current position corresponds to the one recorded by it.

@@

\paragraph*{Pebble pushdown systems}

@@

An {\em $n$-pebble pushdown automaton} 
is a tuple 
$\ppda = (Q, \Gamma,  R, q_{init}, \gamma_{init})$
where $Q$ is a {\em finite set of  states}; $\Gamma$ is a {\em finite stack alphabet};
  $R  \subseteq  Q  \times \Gamma \times
		 \{ 0, 1, \ldots n\} \times \mathcal{P}(\{1,\ldots,n\})
		 \times Q  \times (\Op(\Gamma) \cup \{\text{\textit{drop}}, \text{\textit{lift}}\})$ is a {\em finite set of rules},
		 where the fourth element $S$ of a rule $r$ is a subset of 
		$\mathcal{P}(\{1,\ldots,i\})$ where $i$ is the third element of $r$,
		and if the last element is
		$\text{\textit{lift}}$ then we additionally require $i \in S$;
 $q_{init} \in Q$ is an {\em initial  state};
 $\gamma_{init} \in \Gamma$ is an {\em initial stack symbol}.

@@

The idea of a transition $(q, a, i, S, q', m) \in R$
is that, if the automaton $\ppda$ is in state $q$ with pebbles $1,\ldots, i$ dropped \---- or without pebble dropped if $i = 0$ \---- with top stack symbol $a$, and stack content which corresponds to the stack contents of the pebbles from $S$ and only these pebbles, then
$\ppda$ goes to state 
$q'$ and makes modifications to the stack or the pebbles according to
$m$. 
Note a pebble 
can be lifted only if the stack content which corresponds to the pebble
is the same as the current stack content. 
This is enforced by syntactically requiring
the last pebble dropped $i$ is in the set $S$ used for testing the presence of  pebbles.

@@

An $n$-pebble pushdown automaton $\ppda =  (Q, \Gamma,  R, q_{init}, {\gamma_{init}} )$ induces 
a transition system  $T_{\ppda} = (\Conf(\ppda), \rightarrow_{\ppda})$ 
 such that
for all $(q, a, i, S, q', m) \in R$, with $a \in \Gamma$,
 for all words $w \in  \Gamma^*$, and for all $\{ 1, \ldots , i \}$-pebble assignments $\mu$ such that $\mu(j) = wa$ 
 for each
$j \in S$ 	and
 $\mu(j) \neq wa$ 
 for each $j \in \{ 1, \ldots , i \} \setminus S$,
the following holds
	%\begin{samepage}
\begin{itemize}
\item if $m \in \Op(\Gamma)$, 
either
\begin{itemize}
\item $ m = \text{\textit{push}}^\gamma$ and $(q,wa,\mu) \rightarrow_{\ppda} (q',wa\gamma, \mu)$,

@@

\item $ m = \text{\textit{skip}}$ and $(q,wa,\mu) \rightarrow_{\ppda} (q',wa, \mu)$,
\end{itemize}
\item if $ m = \text{\textit{drop}}$, 
$(q,wa,\mu) \rightarrow_{\ppda} (q',wa, \mu')$,  
where $\mu'$ is the $\{ 1, \ldots , i, i+1 \}$-pebble assignment such that
$\mu'(j) = \mu(j)$, for each
$j \leq i$, and $\mu'(i+1) = wa$, and
\item if $m = \text{\textit{lift}}$, and $i \in S$, i.e. the last pebble dropped belong of the set of pebble we test the presence of,
$(q,wa,\mu) \rightarrow_{\ppda} (q',wa, \mu')$, 
where $\mu'$ is the $\{ 1, \ldots , i - 1 \}$-pebble assignment such that
$\mu'(j) = \mu(j)$, for each
$j < i$.
\end{itemize}
	%\end{samepage}

@@

\problemx{$n$-pebble pushdown parity game}
{An $n$-pebble pushdown automaton $\ppda= (Q, \Gamma,  R, q_{init}, {\gamma_{init}})$, 
where $Q = Q_0  \biguplus Q_1 $, 
and
a priority mapping $\Omega: Q \to \{0, \ldots m\}$.}
{Does player $0$ have a winning strategy from $q_{init}(\gamma_{init}, \mu_{init})$ for the parity game 
$\mathcal{G}= (A_{(\ppda, Q_0 , Q_1)}, \win_{\overline{\Omega}})$ ? \\}

Since pebbles can store stack contents, they can be used to simulate parameters. From this we deduce the following theorem. See Appendix~\ref{appendix A} for more details on the construction.

@@

\paragraph*{Higher-order pushdown systems}

@@

The following operations can be performed on a $1$-stacks of length $m \in \N$.
\begin{eqnarray*}
&\text{\textit{push}}_1^\gamma( w_0 w_1 \ldots w_m) &=\quad w_0 w_1 \ldots w_m \gamma   \text{ for all } \gamma \in  \Gamma, \\
&\text{\textit{pop}}_1(w_0 w_1 \ldots w_{m-1} w_m ) &=\quad w_0 w_1 \ldots w_{m-1}, \text{ if } m \geq 1\\
&\text{\textit{top}}( w_0 w_1 \ldots w_m ) &=\quad w_m.
\end{eqnarray*}
The operations added at level $n+1$ are the copy of the topmost $n$-stack
and the removal of the topmost $n$-stack.
More formally, if $ \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle$ is a stack of 
level $n > 1$, 
the following operations are possible. 
%
\begin{eqnarray*}
&\text{\textit{push}}_n( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &=\quad  \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle \langle s_m \rangle , \\
&\text{\textit{push}}_j( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &=\quad \langle s_0 \rangle, \langle s_1 \rangle \ldots 
\langle \text{\textit{push}}_j(s_m) \rangle,
\text{ if } 2 \leq j < n\\
&\text{\textit{push}}_1^\gamma( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &= \quad
\langle s_0 \rangle \langle s_1 \rangle \ldots 
\langle \text{\textit{push}}_1^\gamma(s_m) \rangle,
\text{ for all } \gamma \in \Gamma,\\
&\text{\textit{pop}}_n( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle \langle s_m \rangle) &= \quad
\langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle
,\\
&\text{\textit{pop}}_j( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle \langle s_m \rangle) &= \quad
\langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle
\langle \text{\textit{pop}}_j(s_m) \rangle,
\text{ if } 1 \leq j < n\\
&\text{\textit{top}}( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &=\quad
\text{\textit{top}}(s_m).
\end{eqnarray*}
The operations $pop_1$ and $top$ are undefined on a stack whose top $1$-stack is empty.

@@

\begin{samepage}
\par\noindent\ignorespacesafterend
A {\em higher-order pushdown systems of level $n$} (or $n$-HPDS for short) 
is a tuple $\H=(Q, \Gamma, R, q_{init}, \gamma_{init})$,
where 
 $Q$ is a non-empty finite {\em set of  states} ;
         $\Gamma$ is a non-empty finite {\em  stack alphabet};
	 $R\subseteq Q\times \Gamma \times Q \times \Op_n(\Gamma)$ is a finite {\em set of  rules};
         $q_{init}$ is the {\em initial  state} ;
         $\gamma_{init} \in \Gamma$ is the {\em %bottom-of-stack
initial stack symbol}.
\end{samepage}

By $\Conf(\mathcal{H}) = Q \times \mathscr{S}_n(\Gamma)$
we denote the set of {\em configurations} of an $n$-HPDS $\H$. 
As usual we abbreviate $(q,s) \in \Conf(\mathcal{H})$ as $q(s)$.
An $n$-HPDS $\H=(Q, \Gamma, R, q_{init})$
induces the transition system $\T_\H = (Conf(\mathcal{H}), \rightarrow_{\mathcal{H}})$
where
for all $q,q'$ in $Q$,
for all $s,s'$ in $\mathscr{S}_n(\Gamma)$,
$q(s) \rightarrow_{\mathcal{H}} q'(s')$ if
there exists some rule
$(q,\gamma,q',op) \in R$
such that
$\text{\textit{top}}(s) = \gamma$
and
$s' = op(s)$.

Again we are interested in parity games.
As expected, given an 
 $n$-HPDS $\mathcal{H}$
 and a partition of
$Q$ into $Q_0$ and $Q_1$,
we partition the configurations of
$T_{\mathcal{H}}$
into
$\Conf_{\mathcal{H},0}=Q_0\times \mathscr{S}_n(\Gamma)$
and
$\Conf_{\mathcal{H},1}=Q_1\times\mathscr{S}_n(\Gamma)$.
%
%
With these notations in mind one can define the arena
$
A_{(\mathcal{H}, Q_0, Q_1)}
=
(\Conf_{\mathcal{H},0}, \Conf_{\mathcal{H},1}, \rightarrow_{\mathcal{H}})$
induced by an $n$-HPDS $\mathcal{H}$ and a partition of its set of states.
